---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


### Loading and preprocessing the data
``` {r preprocessing,echo=TRUE,results="hide" }
library(ggplot2)

#setwd("RDokumente/RepData_PeerAssessment1")
#read in data and get a first overview
activity<-read.csv("activity.csv")
head(activity)
summary(activity)
str(activity)

#recalculate date as date
activity$date<-as.Date(activity$date)

#aggregate steps per interval for average daily activity pattern
stepsPerInterval<-aggregate(steps~interval,data=activity,FUN=mean)
summary(stepsPerInterval)

#aggregate steps per interval for total number of steps taken
stepsPerDay<-aggregate(steps~date,data=activity,FUN=sum)
```

### What is mean total number of steps taken per day?

```{r mean,echo=TRUE}
#Make a histogram of the total number of steps taken each day
plot1<-ggplot(stepsPerDay, aes(x=steps))+geom_histogram(binwidth=100,fill="cornsilk",color="black")
#set the axis labels
plot1+xlab("steps per day")+ylab("no. of days")

#Calculate and report the mean and median total number of steps taken per day
meanSteps<-mean(stepsPerDay$steps)
medianSteps<-median(stepsPerDay$steps)
```

The mean of total number of steps taken per day is `r meanSteps`; the median is `r medianSteps`.

### What is the average daily activity pattern?

``` {r average, echo=TRUE}
# Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
plot2<-ggplot(stepsPerInterval, aes(x=interval,y=steps))+geom_line()
plot2+ylab("average steps during interval")

# Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
maxInterval<-stepsPerInterval$interval[which.max(stepsPerInterval$steps)]
maxSteps<-stepsPerInterval$steps[which.max(stepsPerInterval$steps)]
```

The 5-minute interval `r maxInterval` contains `r maxSteps` steps on average across all the days in the dataset, which is the maximum number of steps.            

### Imputing missing values

``` {r NA, echo=TRUE}
# Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
sumNAs<-sum(is.na(activity))

# Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

# the averages of the intervals calculated for the previous question do replace the missing values 
activityMV <- merge(activity, stepsPerInterval, by = "interval")
mv <- is.na(activityMV$steps.x)
activityMV$steps.x[mv] <- activityMV$steps.y[mv]
activityMV <- activityMV[, c(1:3)]

```

There are `r sumNAs` missing values in the dataset.


### Are there differences in activity patterns between weekdays and weekends?
